---
title: "0. Raw Input Processing"
author: "FOTO NL workflow"
date: "`r format(Sys.Date(), '%Y-%m-%d')`"
output: html_document
---

```{r setup, include=FALSE}
show_script_0_messages <- tolower(Sys.getenv("SCRIPT_0_SHOW_MESSAGES", "false")) == "true"
show_script_0_warnings <- tolower(Sys.getenv("SCRIPT_0_SHOW_WARNINGS", "false")) == "true"
knitr::opts_chunk$set(echo = TRUE, message = show_script_0_messages, warning = show_script_0_warnings)
run_processing <- tolower(Sys.getenv("RUN_0_RAW_INPUT_PROCESSING", "false")) == "true"
```

This stage converts raw chemistry exports into the standardized script_0 input used by the downstream workflow.

The implementation is delegated to the shared reconstruction engine in `helpers/script_0/build_input_common.R` with board definitions from `helpers/script_0/build_input_board_configs.R`. This R Markdown file serves as the documented entry point for script_0 and exposes the runtime switches used for batch runs and debug runs.

At execution time it reads runtime settings from environment variables or command-line arguments, including `RAW_ROOT`, `REFERENCE_INPUT_ROOT`, `OUTPUT_ROOT`, `OUTPUT_FORMAT`, `ONLY_SOURCES`, `OVERWRITE`, `DRY_RUN`, `USE_REFERENCE_FILTER`, `MAX_ROWS_PER_FILE`, `MAX_ROWS_PER_SHEET`, `SKIP_REFERENCE_READ`, `MAX_REFERENCE_ROWS`, and `INCLUDE_LEGACY_INPUTS`. The current `build_*` engine only accepts `OUTPUT_FORMAT=csv`; `TSV_LAYOUT` remains a compatibility variable but is not used by the active writer.

Set `RUN_0_RAW_INPUT_PROCESSING=true` to actually run processing during render.

```{r run-toggle, echo=TRUE}
run_processing
```

```{r script_0_core, eval=run_processing, echo=FALSE, error=FALSE}
script_path <- {
  file_flag <- grep("^--file=", commandArgs(trailingOnly = FALSE), value = TRUE)
  if (length(file_flag) == 0L) "" else sub("^--file=", "", file_flag[1])
}
script_dir <- if (nzchar(script_path)) dirname(normalizePath(script_path)) else getwd()

source(file.path(script_dir, "helpers", "script_0", "build_input_common.R"))
source(file.path(script_dir, "helpers", "script_0", "build_input_board_configs.R"))

run_script_0_pipeline(
  configs = get_board_configs(),
  argv = commandArgs(trailingOnly = TRUE),
  base_dir = script_dir
)
```
