---
title: "script_2_spatial_analysis"
author: "Thomas Hofman"
date: "2025-01-07"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F)
```

## spatial analysis

This script

## Download and install required packages

```{r}
list.of.packages <- c(
  "readr", "dplyr", "tibble", "sf", "mapview", "data.table", "lubridate"
)
missing_pkgs <- setdiff(list.of.packages, rownames(installed.packages()))
if (length(missing_pkgs) > 0) {
  stop(
    "Missing required packages: ",
    paste(missing_pkgs, collapse = ", "),
    ". Install them in bart-env before rendering."
  )
}
suppressPackageStartupMessages(lapply(list.of.packages, require, character.only = TRUE))
```

#### Read the final datafile from script_1\_read_and_clean

extract the coordinates in this file and give a quick overview

```{r include=FALSE}
AllData <- readRDS("ToPAF.rds")

#add year param
AllData$year <- lubridate::year(AllData$SampleDate)

#get unique coords to look up water type
coords <- unique(AllData[,c("SampleID","Locationcode","X", "Y")])
coords_sf <- st_as_sf(coords, coords = c("X", "Y"), crs = "EPSG:28992", remove = F)
coords$xy <- paste0(coords$X,coords$Y)

#quick overview
if (interactive()) {
  mapview(coords_sf, zcol = "SampleID", legend = TRUE, cex = 2)
}
```

#### qgis export

create an export file for further development of maps in QGIS

```{r}
coords_year <- unique(AllData[,c("SampleID","Locationcode","X", "Y", "year")])
write_csv(coords_year, "unique_x_y_mpc.csv")
```

#### extract datapoints outside of NL

Some points are not in NL and are likely to be errors, these are removed from the dataset. As points just outside NL could be real sampling points, a 2km buffer around NL is created.

```{r}
#Read NL shape and convert to simple feature (optional input).
landsgrens_candidates <- c(
  "auxiliary_input/adm_landsgrens.csv",
  "../auxiliary_input/adm_landsgrens.csv",
  "zenodo_ready_20260330/auxiliary_workflow_input/adm_landsgrens.csv",
  "R/adm_landsgrens.csv",
  "R/data/adm_landsgrens.csv",
  "data/adm_landsgrens.csv",
  "data/zenodo_14652471/input/adm_landsgrens.csv"
)
landsgrens_file <- landsgrens_candidates[file.exists(landsgrens_candidates)][1]

if (!is.na(landsgrens_file) && nzchar(landsgrens_file)) {
  NL_shape <- read.csv(landsgrens_file)
  NL_shape_sf <- st_as_sfc(NL_shape$st_astext, crs = "EPSG:28992")
  NL_shape_2023 <- NL_shape_sf[1]

  #buffer two km around NL
  NL_shape_2023_buffer_2k <- sf::st_buffer(NL_shape_2023,2000)

  #get points that are in NL + 2km
  points_in_nl <- suppressWarnings(sf::st_intersection(coords_sf, NL_shape_2023_buffer_2k))

  #get id's of points that are not in NL
  not_nl <- coords_sf[which(!coords_sf$geometry %in% points_in_nl$geometry),]
  not_nl_coords <- do.call(rbind, st_geometry(not_nl)) %>%
    as_tibble() %>% setNames(c("x","y"))

  in_nl <- coords_sf[which(coords_sf$geometry %in% points_in_nl$geometry),]
  in_nl <- do.call(rbind, st_geometry(in_nl)) %>%
    as_tibble() %>% setNames(c("x","y"))
} else {
  warning("adm_landsgrens.csv not found; skipping NL buffer filter and keeping all coordinates.")
  in_nl <- do.call(rbind, st_geometry(coords_sf)) %>%
    as_tibble() %>% setNames(c("x","y"))
  not_nl_coords <- tibble(x = numeric(0), y = numeric(0))
}

#check
if (interactive()) {
  mapview(st_as_sf(in_nl, coords = c("x", "y"), crs = "EPSG:28992"))
}
```

#### Remove coordinates outside the bufferzone and save the output and last clean steps

```{r}
# Format measured values consistently in the final exported CSV.
scientific4 <- function(x) {
  vals <- suppressWarnings(as.numeric(x))
  out <- rep(NA_character_, length(vals))
  ok <- !is.na(vals)
  out[ok] <- sprintf("%.4e", vals[ok])
  out
}

#filter points that are not in NL
AllDataNL <- AllData[which(paste0(AllData$X,"_", AllData$Y) %in% paste0(in_nl$x, "_",in_nl$y)),]


#get nl coords for QGIS maps
coordsNL <- unique(AllDataNL[,c("DataSource","Locationcode","X", "Y")])
write_csv(coordsNL, "coordsNL.csv")

if (interactive()) {
  mapview(st_as_sf(coordsNL, coords = c("X", "Y"), crs = "EPSG:28992"))
}


#clean up 
AllDataNL$year <- lubridate::year(AllDataNL$SampleDate)
AllDataNL$MeasuredValue <- scientific4(AllDataNL$MeasuredValue)
AllDataNL$X <- round(AllDataNL$X, digits = 0)
AllDataNL$Y <- round(AllDataNL$Y, digits = 0)



#save
saveRDS(AllDataNL, "foto_nl_dataset_clean.rds")
if(!dir.exists("output")){dir.create("output")}
fwrite(AllDataNL, "output/foto_nl_dataset_clean.csv", sep = ";")
```
