---
title: "1. combine and clean"
author: "Thomas Hofman"
date: "2025-01-06"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Reading, combining and cleaning of the data

This script reads the separate files of different waterboards and converts them to uniform attributes. This script continues to clean these attributes.

#### install packages

install all required packages and functions

```{r read packages}

# read required packages used by this script
list.of.packages <- c("dplyr")
missing_pkgs <- setdiff(list.of.packages, rownames(installed.packages()))
if (length(missing_pkgs) > 0) {
  stop(
    "Missing required packages: ",
    paste(missing_pkgs, collapse = ", "),
    ". Install them in bart-env before rendering."
  )
}
suppressPackageStartupMessages(lapply(list.of.packages, require, character.only = TRUE))

#load homemade functions required for this script
source("helpers/r/excelDate2Date.R") #used in CheckMapConc function
source("helpers/r/CheckMapConc.R")
source("helpers/r/ValidCAS.R")

#Set folder with script_0 generated input files
#Override with env var PRJDATA_DIR to run from a specific build output.
detect_latest_script_0_root <- function() {
  candidates <- c(
    file.path('..', 'script_0_output'),
    file.path('output', 'script_0_output'),
    file.path('..', 'output', 'script_0_output'),
    list.dirs("output", full.names = TRUE, recursive = FALSE),
    list.dirs("/mnt/scratch_dir/hofmant/biotichs/FOTO_NL_usecase/output", full.names = TRUE, recursive = FALSE)
  )
  candidates <- unique(candidates[grepl("^built_input_from_raw", basename(candidates))])
  if (length(candidates) == 0) return(NA_character_)
  mt <- file.info(candidates)$mtime
  candidates[order(mt, decreasing = TRUE)][1]
}

prjdata <- Sys.getenv("PRJDATA_DIR", unset = "")
if (!nzchar(prjdata)) {
  auto_script_0 <- detect_latest_script_0_root()
  if (!is.na(auto_script_0) && dir.exists(auto_script_0)) {
    prjdata <- auto_script_0
  } else {
    prjdata <- "R/data"
  }
}
input_format <- "csv"
lightweight_checks <- tolower(Sys.getenv("LIGHTWEIGHT_CHECKS", unset = "true")) %in% c("1", "true", "yes")
strict_script_0_diag <- tolower(Sys.getenv("STRICT_SCRIPT_0_DIAG", unset = "false")) %in% c("1", "true", "yes")
message("Using input folder: ", prjdata)
message("Using input format: csv (strict)")
message("Lightweight checks: ", lightweight_checks)
message("Strict script_0 diagnostics: ", strict_script_0_diag)
script_0_board_map <- c(
  "data bestand_Aa en Maas.xlsx" = "aa_en_maas",
  "data bestand_Amstel Gooi Vecht.xlsx" = "amstel_gooi_vecht",
  "data bestand_Brabantse Delta.xlsx" = "brabantse_delta",
  "data bestand_DeDommel.xlsx" = "de_dommel",
  "data bestand_Delfland.xlsx" = "delfland",
  "data bestand_Drents Overijsselse Delta.xlsx" = "drents_overijsselse_delta",
  "data bestand_HH van Rijnland.xlsx" = "hh_van_rijnland",
  "data bestand_Hollands Noorderkwartier.xlsx" = "hollands_noorderkwartier",
  "data bestand_Hollandse Delta.xlsx" = "hollandse_delta",
  "data bestand_Hunze en Aas.xlsx" = "hunze_en_aas",
  "data bestand_Limburg.xlsx" = "limburg",
  "data bestand_WS Noorderzijlvest.xlsx" = "noorderzijlvest",
  "data bestand_Rijn en IJssel.xlsx" = "rijn_en_ijssel",
  "data bestand_Rivierenland.xlsx" = "rivierenland",
  "data bestand_Scheldestromen.xlsx" = "scheldestromen",
  "data bestand_Schieland Krimpenerwaard.xlsx" = "schieland_krimpenerwaard",
  "data bestand_Stichtse Rijnlanden.xlsx" = "stichtse_rijnlanden",
  "data bestand_Vallei Veluwe.xlsx" = "vallei_veluwe",
  "data bestand_Vechtstromen_aangepast 19 aug 2020.xlsx" = "vechtstromen",
  "data bestand_Wetterskip.xlsx" = "wetterskip",
  "data bestand_Zuiderzeeland.xlsx" = "zuiderzeeland"
)

resolve_board_csv <- function(filename) {
  board_id <- unname(script_0_board_map[filename])
  if (length(board_id) == 0L || is.na(board_id) || !nzchar(board_id)) return(NA_character_)
  candidates <- c(
    file.path('..', 'script_0_output'),
    file.path('output', 'script_0_output'),
    file.path('..', 'output', 'script_0_output'),
    file.path(prjdata, "board_input.csv"),
    file.path(prjdata, board_id, "board_input.csv"),
    file.path(prjdata, board_id, paste0(board_id, ".csv"))
  )
  hit <- candidates[file.exists(candidates)]
  if (length(hit) > 0) hit[[1]] else NA_character_
}

input_file <- function(filename) {
  board_csv <- resolve_board_csv(filename)
  if (!is.na(board_csv) && file.exists(board_csv)) return(board_csv)

  stem <- tools::file_path_sans_ext(filename)
  csv_name <- paste0(stem, ".csv")
  search_paths <- c(
    file.path(prjdata, csv_name),
    file.path("R/data", csv_name),
    file.path("auxiliary_input", csv_name),
    file.path("../auxiliary_input", csv_name),
    file.path("data/zenodo_14652471/input", csv_name),
    file.path("data/zenodo_14652471/input/data", csv_name)
  )
  hit <- search_paths[file.exists(search_paths)]
  if (length(hit) > 0) return(hit[[1]])

  original_paths <- c(
    file.path(prjdata, filename),
    file.path("R/data", filename),
    file.path("auxiliary_input", filename),
    file.path("../auxiliary_input", filename),
    file.path("data/zenodo_14652471/input", filename),
    file.path("data/zenodo_14652471/input/data", filename)
  )
  hit_original <- original_paths[file.exists(original_paths)]
  if (length(hit_original) > 0) return(hit_original[[1]])

  stop(
    "CSV input not found for '", filename, "'. ",
    "Expected board_input.csv in script_0 output, ", csv_name, ", or ", filename, " in known data folders."
  )
}

read_input_table <- local({
  function(file, sheet = 1, ...) {
  add_legacy_aliases <- function(df) {
    nms <- names(df)
    # Create dotted aliases for columns with spaces to support legacy mappings
    # used throughout this script (e.g. "Meetpunt.code").
    for (src in nms) {
      dst <- gsub(" +", ".", src)
      if (!identical(dst, src) && !(dst %in% names(df))) {
        df[[dst]] <- df[[src]]
      }
    }
    # Explicit common aliases.
    if ("Meetpunt code" %in% names(df) && !("Meetpunt.code" %in% names(df))) {
      df[["Meetpunt.code"]] <- df[["Meetpunt code"]]
    }
    if ("Meetpunt omschrijving" %in% names(df) && !("Meetpunt.omschrijving" %in% names(df))) {
      df[["Meetpunt.omschrijving"]] <- df[["Meetpunt omschrijving"]]
    }
    df
  }

  if (!file.exists(file)) {
    stop("CSV input not found: ", file)
  }
  ext <- tolower(tools::file_ext(file))
  if (ext %in% c("xlsx", "xls")) {
    if (!requireNamespace("readxl", quietly = TRUE)) {
      stop("Package 'readxl' is required to read Excel inputs: ", file)
    }
    dt <- as.data.frame(readxl::read_excel(file, sheet = sheet))
  } else {
    dt <- read.csv(file, stringsAsFactors = FALSE, check.names = FALSE)
  }
  dt <- add_legacy_aliases(dt)
  dt
}})

check_rows_preserved <- function(board, output_df, ...) {
  input_parts <- list(...)
  input_n <- sum(vapply(input_parts, nrow, integer(1)))
  output_n <- nrow(output_df)
  ok <- identical(output_n, input_n)
  msg <- paste0(
    "Row-preservation check [", board, "]: output=", output_n,
    ", input_sum=", input_n, ", ok=", ok
  )
  if (ok) {
    if (!lightweight_checks) message(msg)
  } else {
    if (lightweight_checks) {
      message(msg)
    } else {
      warning(msg)
    }
  }
  invisible(ok)
}

detect_script_0_root <- function(path) {
  path_norm <- normalizePath(path, mustWork = FALSE)
  if (basename(path_norm) == "data" && dir.exists(dirname(path_norm))) {
    return(dirname(path_norm))
  }
  path_norm
}

read_tsv_safe <- function(path) {
  if (!file.exists(path)) return(NULL)
  read.delim(path, sep = "\t", stringsAsFactors = FALSE, check.names = FALSE)
}

run_script_0_preflight <- function(prjdata, strict = TRUE) {
  script_0_root <- detect_script_0_root(prjdata)
  diag_dir <- file.path(script_0_root, "diagnostics")
  if (!dir.exists(diag_dir)) {
    message("script_0 diagnostics folder not found at: ", diag_dir, " (skipping preflight)")
    return(invisible(NULL))
  }

  diag_files <- list.files(diag_dir, pattern = "_diagnostic\\.tsv$", full.names = TRUE)
  if (length(diag_files) == 0) {
    message("No *_diagnostic.tsv files found in: ", diag_dir)
    return(invisible(NULL))
  }

  diag_list <- lapply(diag_files, read_tsv_safe)
  diag_list <- Filter(Negate(is.null), diag_list)
  diag <- do.call(rbind, diag_list)
  message("Script_0 preflight: found diagnostics for ", nrow(diag), " board(s).")

  required_cols <- c("DataSource", "RawRows", "MappedRows", "FilteredRows", "OutputRows")
  missing_cols <- setdiff(required_cols, names(diag))
  if (length(missing_cols) > 0) {
    warn_msg <- paste0("Script_0 diagnostic columns missing: ", paste(missing_cols, collapse = ", "))
    if (strict) stop(warn_msg) else warning(warn_msg)
  }

  read_error_files <- list.files(diag_dir, pattern = "_read_errors\\.tsv$", full.names = TRUE)
  read_error_rows <- 0L
  if (length(read_error_files) > 0) {
    for (f in read_error_files) {
      dt <- read_tsv_safe(f)
      if (!is.null(dt)) read_error_rows <- read_error_rows + nrow(dt)
    }
  }
  if (read_error_rows > 0L) {
    msg <- paste0("Script_0 preflight detected ", read_error_rows, " read error row(s) in diagnostics.")
    if (strict) stop(msg) else warning(msg)
  } else {
    message("Script_0 preflight: no read errors detected.")
  }

  if (all(c("RawRows", "OutputRows") %in% names(diag))) {
    total_raw <- sum(as.numeric(diag$RawRows), na.rm = TRUE)
    total_out <- sum(as.numeric(diag$OutputRows), na.rm = TRUE)
    message("Script_0 preflight totals: RawRows=", total_raw, ", OutputRows=", total_out)
  }

  invisible(diag)
}

script_0_diagnostics <- run_script_0_preflight(prjdata, strict = strict_script_0_diag)

write_drop_audit <- function(df, path) {
  if (is.null(df) || nrow(df) == 0) return(invisible(NULL))
  write.csv(df, file = path, row.names = FALSE)
  invisible(path)
}

drop_audit_subset <- function(df) {
  keep <- intersect(
    c(
      "Parametercode", "CASnummer", "MeasuredValue", "Unit", "ToUnit",
      "Meetpunt.code", "Meetpunt code", "Datum", "DataSource",
      "PreTreatment", "Limietsymbool"
    ),
    names(df)
  )
  if (length(keep) == 0) return(df[0, , drop = FALSE])
  df[, keep, drop = FALSE]
}

is_concentration_like_unit <- function(unit) {
  unit_norm <- tolower(trimws(as.character(unit)))
  unit_norm[is.na(unit_norm)] <- ""
  grepl("/l$", unit_norm) | grepl("/kg", unit_norm)
}

```

## --------------------------------------------------------------

# Read in and harmonize names datafiles

## --------------------------------------------------------------

Read in script_0 input files and harmonize names with function CheckMapConc

This function also renames the dutch names of variables "OS" and "Corg" into TSS resp. DOC

```{r}
read_board_input <- function(filename, transform_fn = NULL) {
  dt <- read_input_table(input_file(filename))
  if (!is.null(transform_fn)) {
    dt <- transform_fn(dt)
  }
  dt
}

map_board_input <- function(input_dt,
                            data_source,
                            date_as_year_month_day = FALSE,
                            chem_id_names = c("Parametercode", "CASnummer"),
                            x_col = NULL,
                            y_col = NULL) {
  args <- list(
    input_dt,
    SampleID = NA,
    Meetpuntcode = "Meetpunt.code",
    ChemIDnames = chem_id_names,
    MeasuredValue = "Meetwaarde",
    Unit = "Eenheid",
    LimitIndicator = "Limietsymbool",
    SampleDate = "Datum",
    DateAsYearMonthDay = date_as_year_month_day,
    fac2ugL = 1,
    DataSource = data_source
  )
  if (!is.null(x_col)) args$X <- x_col
  if (!is.null(y_col)) args$Y <- y_col
  do.call(CheckMapConc, args)
}

report_board_stats <- function(df) {
  print(length(unique(df$Parametercode)))
  print(length(unique(df$Meetpuntcode)))
  print(summary(df$SampleDate))
}

check_and_report_board <- function(board, output_df, input_dt) {
  check_rows_preserved(board, output_df, input_dt)
  report_board_stats(output_df)
}
```

### Read and harmonize all board inputs

The expensive board read/map step can run in parallel. Set `STAGE1_N_CORES` to opt into multicore execution; default is `1` to keep behavior deterministic outside managed runs.

```{r}
Num2Date <- function (x) {
  out <- as.character(x)
  out[is.na(out)] <- NA_character_
  out <- trimws(out)

  # Keep script_0 ISO dates unchanged; only rewrite legacy compact YYYYMMDD values.
  compact <- grepl("^[0-9]{8}$", out)
  out[compact] <- paste(substr(out[compact], 1, 4), substr(out[compact], 5, 6), substr(out[compact], 7, 8), sep = "-")
  out
}

stage1_n_cores <- suppressWarnings(as.integer(Sys.getenv("STAGE1_N_CORES", unset = "1")))
if (is.na(stage1_n_cores) || stage1_n_cores < 1L) stage1_n_cores <- 1L
message("Stage-1 board read/map cores: ", stage1_n_cores)

run_stage1_lapply <- function(X, FUN, n_cores = 1L) {
  if (.Platform$OS.type == "unix" && n_cores > 1L) {
    parallel::mclapply(X, FUN, mc.cores = n_cores, mc.preschedule = FALSE)
  } else {
    lapply(X, FUN)
  }
}

apply_drents_location_files <- function(dt) {
  loc_file_1 <- input_file("AanvlocDrent.xlsx")
  if (file.exists(loc_file_1)) {
    LocDrent <- read_input_table(loc_file_1)
    LocDrent <- LocDrent[!is.na(LocDrent$GeometriePunt.X_RD),]
    matchLoc <- match(dt$Meetpuntcode, LocDrent$Identificatie)
    dt$X <- LocDrent$GeometriePunt.X_RD[matchLoc]
    dt$Y <- LocDrent$GeometriePunt.Y_RD[matchLoc]
  }
  loc_file_2 <- input_file("lokatie info_Drents Overijsselse Delta.xlsx")
  if (file.exists(loc_file_2)) {
    LocDrent <- read_input_table(loc_file_2)
    LocDrent <- LocDrent[!is.na(LocDrent$GeometriePunt.X_RD),]
    matchLoc <- match(dt$Meetpuntcode, LocDrent$Identificatie)
    dt$X <- LocDrent$GeometriePunt.X_RD[matchLoc]
    dt$Y <- LocDrent$GeometriePunt.Y_RD[matchLoc]
  }
  dt
}

build_board_dataset <- function(spec) {
  input_dt <- read_board_input(spec$filename, transform_fn = spec$transform_fn)
  map_args <- c(
    list(
      input_dt = input_dt,
      data_source = spec$data_source
    ),
    spec$map_args
  )
  output_dt <- do.call(map_board_input, map_args)
  if (!is.null(spec$postprocess_fn)) {
    output_dt <- spec$postprocess_fn(output_dt, input_dt)
  }
  list(
    input = input_dt,
    output = output_dt,
    board = spec$board_label
  )
}

board_specs <- list(
  list(var_name = "AaenMaas", input_name = "AaenMaas_input", filename = "data bestand_Aa en Maas.xlsx", data_source = "Aa en Maas", board_label = "Aa en Maas", map_args = list(), transform_fn = NULL, postprocess_fn = NULL),
  list(var_name = "AmstelGooiVecht", input_name = "AmstelGooiVecht_input", filename = "data bestand_Amstel Gooi Vecht.xlsx", data_source = "Amstel Gooi Vecht", board_label = "Amstel Gooi Vecht", map_args = list(), transform_fn = NULL, postprocess_fn = NULL),
  list(var_name = "BrabantseDelta", input_name = "BrabantseDelta_input", filename = "data bestand_Brabantse Delta.xlsx", data_source = "Brabantse Delta", board_label = "Brabantse Delta", map_args = list(), transform_fn = NULL, postprocess_fn = function(dt, input_dt) dt[dt$X < 400000, ]),
  list(var_name = "DeDommel", input_name = "DeDommel_input", filename = "data bestand_DeDommel.xlsx", data_source = "De Dommel", board_label = "De Dommel", map_args = list(), transform_fn = NULL, postprocess_fn = NULL),
  list(var_name = "Delfland", input_name = "Delfland_input", filename = "data bestand_Delfland.xlsx", data_source = "Delfland", board_label = "Delfland", map_args = list(), transform_fn = NULL, postprocess_fn = NULL),
  list(var_name = "DrentsOverijsselseDelta", input_name = "DrentsOverijsselseDelta_input", filename = "data bestand_Drents Overijsselse Delta.xlsx", data_source = "Drents Overijsselse Delta", board_label = "Drents Overijsselse Delta", map_args = list(), transform_fn = NULL, postprocess_fn = function(dt, input_dt) apply_drents_location_files(dt)),
  list(var_name = "HHvanRijnland", input_name = "HHvanRijnland_input", filename = "data bestand_HH van Rijnland.xlsx", data_source = "HH van Rijnland", board_label = "HH van Rijnland", map_args = list(date_as_year_month_day = "%Y-%m-%d"), transform_fn = function(dt) { dt$Datum <- Num2Date(dt$Datum); dt }, postprocess_fn = NULL),
  list(var_name = "HollandsNoorderkwartier", input_name = "HollandsNoorderkwartier_input", filename = "data bestand_Hollands Noorderkwartier.xlsx", data_source = "Hollands Noorderkwartier", board_label = "Hollands Noorderkwartier", map_args = list(), transform_fn = NULL, postprocess_fn = NULL),
  list(var_name = "HollandseDelta", input_name = "HollandseDelta_input", filename = "data bestand_Hollandse Delta.xlsx", data_source = "Hollandse Delta", board_label = "Hollandse Delta", map_args = list(), transform_fn = NULL, postprocess_fn = NULL),
  list(var_name = "HunzeenAas", input_name = "HunzeenAas_input", filename = "data bestand_Hunze en Aas.xlsx", data_source = "Hunze en Aas", board_label = "Hunze en Aas", map_args = list(), transform_fn = NULL, postprocess_fn = NULL),
  list(var_name = "Limburg", input_name = "Limburg_input", filename = "data bestand_Limburg.xlsx", data_source = "Limburg", board_label = "Limburg", map_args = list(), transform_fn = NULL, postprocess_fn = NULL),
  list(var_name = "RijnenIJssel", input_name = "RijnenIJssel_input", filename = "data bestand_Rijn en IJssel.xlsx", data_source = "Rijn en IJssel", board_label = "Rijn en IJssel", map_args = list(), transform_fn = NULL, postprocess_fn = NULL),
  list(var_name = "Rivierenland", input_name = "Rivierenland_input", filename = "data bestand_Rivierenland.xlsx", data_source = "Rivierenland", board_label = "Rivierenland", map_args = list(), transform_fn = NULL, postprocess_fn = NULL),
  list(var_name = "Scheldestromen", input_name = "Scheldestromen_input", filename = "data bestand_Scheldestromen.xlsx", data_source = "Scheldestromen", board_label = "Scheldestromen", map_args = list(), transform_fn = NULL, postprocess_fn = NULL),
  list(var_name = "SchielandKrimpenerwaard", input_name = "SchielandKrimpenerwaard_input", filename = "data bestand_Schieland Krimpenerwaard.xlsx", data_source = "Schieland Krimpenerwaard", board_label = "Schieland Krimpenerwaard", map_args = list(), transform_fn = NULL, postprocess_fn = NULL),
  list(var_name = "StichtseRijnlanden", input_name = "StichtseRijnlanden_input", filename = "data bestand_Stichtse Rijnlanden.xlsx", data_source = "Stichtse Rijnlanden", board_label = "Stichtse Rijnlanden", map_args = list(), transform_fn = NULL, postprocess_fn = NULL),
  list(var_name = "ValleiVeluwe", input_name = "ValleiVeluwe_input", filename = "data bestand_Vallei Veluwe.xlsx", data_source = "Vallei Veluwe", board_label = "Vallei Veluwe", map_args = list(), transform_fn = NULL, postprocess_fn = NULL),
  list(var_name = "Vechtstromen", input_name = "Vechtstromen_input", filename = "data bestand_Vechtstromen_aangepast 19 aug 2020.xlsx", data_source = "Vechtstromen", board_label = "Vechtstromen", map_args = list(), transform_fn = NULL, postprocess_fn = NULL),
  list(var_name = "Wetterskip", input_name = "Wetterskip_input", filename = "data bestand_Wetterskip.xlsx", data_source = "Wetterskip", board_label = "Wetterskip", map_args = list(), transform_fn = NULL, postprocess_fn = NULL),
  list(var_name = "Noorderzijlvest", input_name = "Noorderzijlvest_input", filename = "data bestand_WS Noorderzijlvest.xlsx", data_source = "Noorderzijlvest", board_label = "Noorderzijlvest", map_args = list(date_as_year_month_day = "%Y-%m-%d", x_col = "xcoordinaat", y_col = "ycoordinaat"), transform_fn = NULL, postprocess_fn = NULL),
  list(var_name = "Zuiderzeeland", input_name = "Zuiderzeeland_input", filename = "data bestand_Zuiderzeeland.xlsx", data_source = "Zuiderzeeland", board_label = "Zuiderzeeland", map_args = list(), transform_fn = NULL, postprocess_fn = NULL),
  list(var_name = "Maas", input_name = "Maas_input", filename = "Rijkswaterstaat data bestand_Maas.xlsx", data_source = "Maas", board_label = "Rijkswaterstaat Maas", map_args = list(date_as_year_month_day = "%d-%m-%Y", chem_id_names = c("Parametercode", "Casnummer")), transform_fn = NULL, postprocess_fn = function(dt, input_dt) { names(dt)[names(dt) == "Casnummer"] <- "CASnummer"; dt$X <- dt$X / 100; dt$Y <- dt$Y / 100; dt }),
  list(var_name = "Rijn", input_name = "Rijn_input", filename = "Rijkswaterstaat data bestand_Rijn.xlsx", data_source = "Rijn", board_label = "Rijkswaterstaat Rijn", map_args = list(date_as_year_month_day = "%d-%m-%Y", chem_id_names = c("Parametercode", "Casnummer")), transform_fn = NULL, postprocess_fn = function(dt, input_dt) { names(dt)[names(dt) == "Casnummer"] <- "CASnummer"; dt$X <- dt$X / 100; dt$Y <- dt$Y / 100; dt })
)

board_results <- run_stage1_lapply(board_specs, build_board_dataset, n_cores = stage1_n_cores)
target_env <- environment()
for (i in seq_along(board_specs)) {
  spec <- board_specs[[i]]
  res <- board_results[[i]]
  assign(spec$input_name, res$input, envir = target_env)
  assign(spec$var_name, res$output, envir = target_env)
  check_and_report_board(spec$board_label, res$output, res$input)
}
rm(board_results, board_specs, target_env)
```


### Binding all files into one dataframe

bind all dataframes to one single dataframe and remove seperate frames

```{r}
AllData8 <- rbind(AaenMaas, AmstelGooiVecht, BrabantseDelta, DeDommel, Delfland,
                  DrentsOverijsselseDelta, HHvanRijnland, HollandseDelta, HollandsNoorderkwartier,
                  HunzeenAas, Limburg, Noorderzijlvest, Maas, Rijn, RijnenIJssel, Rivierenland, Scheldestromen,
                  SchielandKrimpenerwaard, StichtseRijnlanden, ValleiVeluwe, Vechtstromen,
                  Wetterskip, Zuiderzeeland)
rm(AaenMaas, AmstelGooiVecht, BrabantseDelta, DeDommel, Delfland,
   DrentsOverijsselseDelta, HHvanRijnland, HollandseDelta, HollandsNoorderkwartier,
   HunzeenAas, Limburg, Noorderzijlvest, Maas, Rijn, RijnenIJssel, Rivierenland, Scheldestromen,
   SchielandKrimpenerwaard, StichtseRijnlanden, ValleiVeluwe, Vechtstromen,
   Wetterskip, Zuiderzeeland)

# Robust fallback: if DataSource is empty, recover from Waterschap label.
idx_missing_ds <- is.na(AllData8$DataSource) | trimws(as.character(AllData8$DataSource)) == ""
idx_has_ws <- !is.na(AllData8$Waterschap) & trimws(as.character(AllData8$Waterschap)) != ""
idx_fill_ds <- idx_missing_ds & idx_has_ws
if (any(idx_fill_ds)) {
  AllData8$DataSource[idx_fill_ds] <- as.character(AllData8$Waterschap[idx_fill_ds])
}
```

## Checkpoint 1: InBAllData.RDS

in between save of complete single dataframe with all waterboards in it

```{r}
saveRDS(AllData8, file = "InBAllData.RDS")
```

## --------------------------------------------------------------

# Clean coordinates and and unique location codes

## --------------------------------------------------------------

*if data not active run: AllData8 \<- readRDS(file = "InBAllData.RDS")*

### Assign letter to each waterboard and add this to locationcode ("Meetpuntcode")

Some location codes are identical across water boards: add a unique letter per water board to Meetpuntcode (location code) to make every Meetpuntcode unique

```{r}
WSen <- data.frame(
  WS = unique(AllData8$DataSource), stringsAsFactors = F
)
WSen$Char <- LETTERS[1:nrow(WSen)]
#add assigned letter of water board to location code ("Meetpuntcode")
for (WS in 1:nrow(WSen)){
  ws_name <- WSen$WS[WS]
  if (is.na(ws_name)) next
  idx_ws <- !is.na(AllData8$DataSource) & AllData8$DataSource == ws_name
  if (any(idx_ws)) {
    mp <- as.character(AllData8$Meetpuntcode[idx_ws])
    AllData8$Meetpuntcode[idx_ws] <-
      ifelse(is.na(mp), NA_character_, paste(WSen$Char[WS], mp, sep = ""))
  }
}
```

### Clean up location-XY relation there are XY with multiple locationcode and locationcode with multiple XY

```{r}
#set X and Y as numeric
AllData8$X <- as.numeric(AllData8$X)
AllData8$Y <- as.numeric(AllData8$Y)

#Small X,Y deviations for the same location("Meetpuntcode")
#correct to only one set of X and Y per location -> use most used X and Y
uniqLocations <- unique(AllData8[,c("Meetpuntcode","X","Y")])
tabUniqLoc <- table (uniqLocations$Meetpuntcode) #15 locations ("Meetpuntcode") identified with multiple x and y coordinates
multi_names <- names(tabUniqLoc[tabUniqLoc > 1])
MultiMeetpunt <- data.frame(
  MultiMeetpuntName = multi_names,
  mostX = rep(NA_character_, length(multi_names)),
  mostY = rep(NA_character_, length(multi_names)),
  stringsAsFactors = F
)
for (imeetpunt in seq_len(nrow(MultiMeetpunt))) {#imeetpunt = 2
  tabInAll <- table(AllData8[AllData8$Meetpuntcode == MultiMeetpunt$MultiMeetpuntName[imeetpunt], c("X","Y")])
  XYmax <- which.max(tabInAll) -1
  Xmax <- 1 + XYmax %% 2
  Ymax <- 1 + XYmax %/% 2
  MultiMeetpunt$mostX[imeetpunt] <- rownames(tabInAll)[Xmax]
  MultiMeetpunt$mostY[imeetpunt] <- colnames(tabInAll)[Ymax]
  AllData8$X[AllData8$Meetpuntcode == MultiMeetpunt$MultiMeetpuntName[imeetpunt]] <- as.numeric(MultiMeetpunt$mostX[imeetpunt])
  AllData8$Y[AllData8$Meetpuntcode == MultiMeetpunt$MultiMeetpuntName[imeetpunt]] <- as.numeric(MultiMeetpunt$mostY[imeetpunt])
}

#test on success of clean-up of location data
wasNum <- nrow(uniqLocations)
uniqLocations <- unique(AllData8[,c("Meetpuntcode","X","Y")])
stopifnot(wasNum - nrow(MultiMeetpunt) == nrow(uniqLocations))

# Strict coordinate requirement: keep only rows with valid X and Y.
AllData8 <- AllData8[
  !is.na(AllData8$X) & AllData8$X > 0 &
  !is.na(AllData8$Y) & AllData8$Y > 0,
]

```

### X and Y with multiple assigned location codes (n = 354)

Correct to only one location code per X,Y combination -\> choose most used

```{r}
AggXY <- aggregate(Meetpuntcode ~ X + Y, data = uniqLocations, FUN = length)
MultiMeetpunt = AggXY[AggXY$Meetpuntcode > 1, c("X","Y")]
for (imeetpunt in seq_len(nrow(MultiMeetpunt))) {#imeetpunt = 2
  PrepTable <- AllData8[AllData8$X == MultiMeetpunt$X[imeetpunt] & AllData8$Y == MultiMeetpunt$Y[imeetpunt],c("X","Y","Meetpuntcode")]
  Ltable <- table(PrepTable$Meetpuntcode)
  mostFreq <- names(Ltable)[which.max(Ltable)]
  AllData8$Meetpuntcode[AllData8$X == MultiMeetpunt$X[imeetpunt] & AllData8$Y == MultiMeetpunt$Y[imeetpunt]] <- mostFreq
}
```

### sometimes X and Y are in km instead of m

correcting units from km to m

```{r}
idx_x_km <- !is.na(AllData8$X) & AllData8$X < 1000
idx_y_km <- !is.na(AllData8$Y) & AllData8$Y < 1000
AllData8$X[idx_x_km] <- AllData8$X[idx_x_km] * 1000
AllData8$Y[idx_y_km] <- AllData8$Y[idx_y_km] * 1000
```

## --------------------------------------------------------------

# Cleaning substance descriptions

## --------------------------------------------------------------

in 5 steps:

-   1\. Remove measurements with no substance descriptors (Parametercode and CAS)

-   2\. Add missing CAS numbers based on substance names (Parametercode)

-   3\. Check for invalid CAS numbers based on their checksum

-   4\. Assign a preferred unit to valid CAS numbers, as well as modifying factors

-   5\. Remove all measurements/substances with invalid CAS numbers, except for modifying factors

### Overview of NA parametercodes

```{r}
table(is.na(AllData8$Parametercode)) #n = 23385 NA
```

### remove all data with NA Parametercodes

These do also not have a CAS, no way to identify substance

```{r}
AllData8 <- AllData8[!is.na(AllData8$Parametercode),]
```

Substances are identified by either CAS number or Aquocode ("Parametercode")

Add missing CASnumbers from AquoParams dataset based on Aquocode ("Parametercode") (n= 2999 paramatercodes with missing CAS number)

```{r}
load("resources/support_tables/AquoParams.rda")
matchAquo <- match(AllData8$Parametercode,AquoParams$Code)
AllData8$CASnummer[is.na(AllData8$CASnummer)] <- AquoParams$CASnummer[matchAquo[is.na(AllData8$CASnummer)]]
NoParameter <- AllData8[is.na(matchAquo)&is.na(AllData8$CASnummer),] # none
#Check amount of unique parameters and amount of unique CAS numbers
length(unique(AllData8$Parametercode)) #n = 1830
length(unique(AllData8$CASnummer)) #n = 1681 -> not all measured substances/chemicals/units have a CAS code, such as Temperature or pH
```

#### *Intermezzo: Name and CAS of conductivity needs to be changed here, otherwise these measurements are removed in these steps ----*

##### remove where Rgeleidbaarheid and Unit == ug/l (n = 239)

```{r}
AllData8 <- AllData8[!(AllData8$Parametercode == "GELDHD" & AllData8$Unit == "ug/l"),] # n = 239
Rgeleidbaarheid <- AllData8$Parametercode == "GELDHD"
#Inidcate unit for conductivity ("Rgeleidbaarheid") is mS/m for water board "Drents Overijsselse Delta"
AllData8$Unit[Rgeleidbaarheid & AllData8$DataSource == "Drents Overijsselse Delta"] <- "mS/m"
AllData8$Parametercode[Rgeleidbaarheid] <- "Cl"
AllData8$CASnummer[Rgeleidbaarheid] <-  "16887-00-6"
```

#### *Continue substance description cleaning*

#### define table with all unique measured variables and corresponding CAS numbers

```{r}

VarUnit <- unique(AllData8[,c("Parametercode","CASnummer")])
#Check for invalid CAS numbers using ValidCAS function
#Checks the validity of a CAS code based on the checksum of CAS numbers
VarUnit$ValidCAS <- ValidCAS(VarUnit$CASnummer) # n = 147 invalid CAS numbers
table(VarUnit$ValidCAS)

#Create table with names and units of modifiers from "ModifierDefaults" dataset
#For later use: to differentiate modifying factors(for bioavailability) from contaminating substances
load("resources/support_tables/ModifierDefaults.rda")
MODnameUnit <- sapply(names(ModifierDefaults[-1]), strsplit, split="\\(")
MODname <- sapply(MODnameUnit, function(x) x[1])
MODnameUnit <- tolower(sapply(MODnameUnit, function(x) substr(x[2],start = 1,stop = nchar(x[2])-1)))
names(MODnameUnit) <- MODname

#Correct parametercode T to the more correct Tw (Temperature of water instead of just temperature)
AllData8$Parametercode[AllData8$Parametercode == "T"] <- "Tw"

#Add logical variable to dataframe indicating wether CAS is valid (TRUE) or not (FALSE)
UniqCASinData <- unique(AllData8$CASnummer)

valid_cas_flags <- ValidCAS(UniqCASinData) #using ValidCAS function to check if CAS numbers are valid
UniqCASinData[!valid_cas_flags] #invalid CAS: "NVT" (Dutch acronym for not applicable), and "som van stoffen" (Dutch for sum of substances)
UniqCASinData <- UniqCASinData[valid_cas_flags]

AllData8$ValidCAS <- AllData8$CASnummer %in% UniqCASinData

#Indicate that measurements with validCAS = TRUE have preferred unit of "ug/l"
AllData8$ToUnit <- ""
AllData8$ToUnit[AllData8$ValidCAS] <- "ug/l"

#Retrieve preferred unit for modifying factors from MODname
for (Modfy in MODname) {
  AllData8$ToUnit[AllData8$Parametercode == Modfy] <- MODnameUnit[Modfy]
}
```

#### Checkpoint 2: AllData.RDS

```{r}
#save checkpoint data
saveRDS(AllData8, file = "AllData.RDS")

#write VarUnit for sup info report 
readr::write_delim(VarUnit, "unique_valid_CAS_codes.csv", delim = ";")
```

#### Create df OthParam: measurements with an invalid CASnr and are also not modifying factors

filter based on available ToUnit, as parameters with valid CAS or modifying factors have a unit indicated in ToUnit

```{r}
OthParam <- AllData8[AllData8$ToUnit == "",] # n = 8896696

OthParam$Valid <- !is.na(OthParam$MeasuredValue) & OthParam$LimitIndicator == "-"
othervalid <- aggregate(MeasuredValue~Parametercode + DataSource, data = OthParam[OthParam$Valid==T,], FUN = length) #Df with amount of measurements per parameter en datasource
write.csv(othervalid, file = "othervalid.csv")
length(unique(OthParam$Parametercode)) #143
```

We don't want to throw away all measurement in OthParam (unique n=144): some are relevant ecological paramaters such as turbidity and sums of substances \#### Load in xlsx file indicating wether to keep a parametercode or not These were checked manually based on their Aquocode. Measurements in OthParam that were unclear were marked to be thrown away

```{r}
SI_2_table <- read_input_table(input_file("SI_2_table.xlsx"))


#### Adding some missing CAS numbers of substances
unique(AllData8$CASnummer[AllData8$Parametercode == "DccC6ySn"])
ValidCAS("7529-52-4")
AllData8$CASnummer[AllData8$Parametercode == "DccC6ySn"] <- "7529-52-4"

unique(AllData8$CASnummer[AllData8$Parametercode == "DFySn"])
ValidCAS("1011-95-6")
AllData8$CASnummer[AllData8$Parametercode == "DFySn"] <- "7529-52-4"

unique(AllData8$CASnummer[AllData8$Parametercode == "FEOa"])
ValidCAS("603-17-8")
AllData8$CASnummer[AllData8$Parametercode == "FEOa"] <- "603-17-8"

unique(AllData8$CASnummer[AllData8$Parametercode == "TccC6ySn"])
ValidCAS("13121-70-5")
AllData8$CASnummer[AllData8$Parametercode == "TccC6ySn"] <- "13121-70-5"

#Check if there are any Parametercodes in OthParam that are not in SI_2_table
OthParamUniq <- unique(OthParam$Parametercode)
OthParamUniq[!OthParamUniq %in% SI_2_table$Parametercode] #check if all parametercodes in OthParam are listed in SI_2_table -> they are
rm(OthParamUniq)
SI_2_table <- mutate(SI_2_table, Keep = as.logical(Keep)) #Making sure the colomn indicating whether to keep a measurement is a logical
table(SI_2_table$Keep) #4 true
GoodToGo <- filter(SI_2_table, Keep == FALSE) #Df with all Parametercodes that can be discarded

#Filter AllData8 based on df GoodToGo
AllData8$Keep <- !AllData8$Parametercode %in% GoodToGo$Parametercode
table(AllData8$Keep) #n=8481632 measurements discarded
#Check if AllData$Keep completely corresponds with GoodToGo
GoodToGoCheck <- filter(AllData8, Keep == FALSE)
GoodToGoCheck <- unique(GoodToGoCheck$Parametercode)
GoodToGoCheck[GoodToGoCheck %in% GoodToGo$Parametercode]
```

#### Create df ToPAF from AllData8, only measurements that we want to keep.

Data we want to keep are measurements with valid cas codes, measurements of modifiers. Other parameters, such as turbidity, organic N, water depth, sums of substances, and unclear parameters, are discarded as indicated in the SI_2_table file

```{r}
ToPAF <- filter(AllData8, Keep ==TRUE)
save(ToPAF, file = "ToPAF.rda")
```

Remove df "AllData8" , rest of script works with new df "ToPAF"

```{r}
rm(AllData8)
```

## --------------------------------------------------------------

# Unit and value conversion and addition

## --------------------------------------------------------------

#### Add molecular weight to dataframe

molecular weight is later used in unit conversion section

```{r}
#Add column with molecular weight values to df "ToPAF" based on name match with Gross dataset
load("resources/support_tables/Gross.rda")
MatchDef <- match(ToPAF$CASnummer, Gross$CAS)
MatchDef[is.na(MatchDef)] <- match(ToPAF$Parametercode[is.na(MatchDef)], Gross$AquoCode)
ToPAF$MolM <- Gross$MW.g.Mol[MatchDef]

#Add molecular weight values for modifying factors based on name match with "Gross" dataset
matchAquoGross <- match(MODname,Gross$AquoCode)
for (matchMod in MODname[!is.na(matchAquoGross)]){ #matchMod = "Ca"
  ToPAF$MolM[ToPAF$Parametercode==matchMod] <- Gross$MW.g.Mol[match(matchMod, Gross$AquoCode)]
}
```

#### Curation of "PreTreatment" variable

Describes properties (such as measurered after filtering) of observations

```{r}
PropertyCodes <- unique(ToPAF$PreTreatment)
list_PropertyCodes <- list()
for (Property in PropertyCodes){
  #Property <- "nf"
  slice <- filter(ToPAF, PreTreatment == Property)
  slicetable <- data.frame(table(slice$Parametercode))

  list_PropertyCodes[[length(list_PropertyCodes) + 1]] <- slicetable
}
names(list_PropertyCodes) <- PropertyCodes
```

#### Unit and Value conversions

```{r}
#unit for pH not always described as "pH". We assumed used unit is pH anyway as no other units are possible/commonly used
#Set unit of pH to same unit as used in "ModifierDefaults" dataset (which is "units")

ToPAF$Unit[ToPAF$Parametercode=="pH"] <- MODnameUnit["pH"]

#Measurements selected in SI_2_table do not have preferred unit indicated yet in ToUnit column
#Set ToUnit in this section, for later unit conversion

ToPAF$ToUnit[ToPAF$Parametercode == "FEOa"] <- "ug/l"
ToPAF$ToUnit[ToPAF$Parametercode == "DFySn"] <- "ug/l"
ToPAF$ToUnit[ToPAF$Parametercode == "TccC6ySn"] <- "ug/l"

# O2 is reported both as concentration and saturation.
# Keep saturation values in %, but normalize concentration-style rows to ug/l so
# they no longer survive as unresolved mg/l mismatches.
o2_pct <- ToPAF$Parametercode == "O2" & ToPAF$Unit == "%"
ToPAF$ToUnit[o2_pct] <- "%"
o2_conc <- ToPAF$Parametercode == "O2" & ToPAF$Unit %in% c("ug/l", "mg/l")
ToPAF$ToUnit[o2_conc] <- "ug/l"
rm(o2_pct, o2_conc)

table(ToPAF$ToUnit)

#Convert measurement of data to their preferred unit (as listed in column "ToUnit"), where needed
#For measurements with measurement in mg/L but preferred unit in microg/L:
mg2ug <- ToPAF$Unit == "mg/l" & ToPAF$ToUnit == "ug/l"
ToPAF$MeasuredValue[mg2ug] <- ToPAF$MeasuredValue[mg2ug] * 1000
ToPAF$Unit[mg2ug] <- "ug/l"
rm(mg2ug)

#For measurements with measurement in microg/L but preferred unit in mg/L:
#This is the case for (some measurements of) modifying factors DOC and TSS
ug2mg <- ToPAF$Unit == "ug/l" & ToPAF$ToUnit == "mg/l"
ToPAF$MeasuredValue[ug2mg] <- ToPAF$MeasuredValue[ug2mg] / 1000
ToPAF$Unit[ug2mg] <- "mg/l"
rm(ug2mg)

#For measurements with measurement in ng/L but preferred unit in microg/L:
ng2ug <- ToPAF$Unit == "ng/l" & ToPAF$ToUnit == "ug/l"
ToPAF$MeasuredValue[ng2ug] <- ToPAF$MeasuredValue[ng2ug] / 1000
ToPAF$Unit[ng2ug] <- "ug/l"
rm(ng2ug)

#For measurements with measurement in pg/L but preferred unit in microg/L:
pg2ug <- ToPAF$Unit == "pg/l" & ToPAF$ToUnit == "ug/l"
ToPAF$MeasuredValue[pg2ug] <- ToPAF$MeasuredValue[pg2ug] / 1000000
ToPAF$Unit[pg2ug] <- "ug/l"
rm(pg2ug)

#For measurements with measurement in mmol/L but preferred unit in microg/L:
mmolL2ugL <- ToPAF$Unit == "mmol/l" & ToPAF$ToUnit == "ug/l"
ToPAF$MeasuredValue[mmolL2ugL] <- ToPAF$MeasuredValue[mmolL2ugL] * ToPAF$MolM[mmolL2ugL] * 1000
ToPAF$Unit[mmolL2ugL] <- "ug/l"
rm(mmolL2ugL)

#Cl includes both concentrations and conductivity measurements
#Converting conductivity to Cl concentration gives strange results (see section below)
#Give conductivity it's own parametercode and harmonize use of unit to "mS/cm"
table(ToPAF$Unit[ToPAF$Parametercode == "Cl"])
ToPAF$Parametercode[ToPAF$Parametercode == "Cl" & ToPAF$Unit != "ug/l"] <- "Conductivity"
table(ToPAF$Unit[ToPAF$Parametercode == "Conductivity"])

ToPAF$ToUnit[ToPAF$Parametercode == "Conductivity"] <-  "mS/cm"
ToPAF$ValidCAS[ToPAF$Parametercode == "Conductivity"] <- F
ToPAF$CASnummer[ToPAF$Parametercode == "Conductivity"] <- "NVT"

mSm2mScm <- ToPAF$Unit == "mS/m" & ToPAF$ToUnit == "mS/cm"
ToPAF$MeasuredValue[mSm2mScm] <- ToPAF$MeasuredValue[mSm2mScm] / 100
ToPAF$Unit[mSm2mScm] <- "mS/cm"
rm(mSm2mScm)

uScm2mScm <- ToPAF$Unit == "uS/cm" & ToPAF$ToUnit == "mS/cm"
ToPAF$MeasuredValue[uScm2mScm] <- ToPAF$MeasuredValue[uScm2mScm] / 1000
ToPAF$Unit[uScm2mScm] <- "mS/cm"
rm(uScm2mScm)


#Some NOxHy is given as N or Nnf;  these need converting to actual ug by:
#Molweight N = 14, H = 1, O = 16

ToPAF$MeasuredValue[ToPAF$Parametercode=="NH4" & ToPAF$PreTreatment %in% c("N","Nnf")] <-
    ToPAF$MeasuredValue[ToPAF$Parametercode=="NH4" & ToPAF$PreTreatment %in% c("N","Nnf")] * 18 / 14
ToPAF$MeasuredValue[ToPAF$Parametercode=="NH3" & ToPAF$PreTreatment %in% c("N","Nnf")] <-
  ToPAF$MeasuredValue[ToPAF$Parametercode=="NH3" & ToPAF$PreTreatment %in% c("N","Nnf")] * 17 / 14
ToPAF$MeasuredValue[ToPAF$Parametercode=="NO3" & ToPAF$PreTreatment %in% c("N","Nnf")] <-
  ToPAF$MeasuredValue[ToPAF$Parametercode=="NO3" & ToPAF$PreTreatment %in% c("N","Nnf")] * (14+3*16) / 14
ToPAF$MeasuredValue[ToPAF$Parametercode=="NO2" & ToPAF$PreTreatment %in% c("N","Nnf")] <-
  ToPAF$MeasuredValue[ToPAF$Parametercode=="NO2" & ToPAF$PreTreatment %in% c("N","Nnf")] * (14+2*16) / 14

#Check if all units are now identical to their intended unit
Unitmm <- unique(ToPAF[ToPAF$Unit!=ToPAF$ToUnit,c("Parametercode" ,"Unit", "ToUnit", "MolM")])

save(ToPAF, file = "ToPAF.rda")
```

#### Some mismatches remain, resolved below:

```{r}
#Some units for temperature (Tw) are in ug/l instead of oC
#create seperate object of Tw measurements to take a closer look at errors
ToPAFTw <- filter(ToPAF, Parametercode == "Tw")
table(ToPAFTw$Unit) #224 measurements with wrong unit of ug/l
#create seperate object of Tw with wrong unit to take a closer look
ToPAFtwmicrogram <- filter(ToPAFTw, Unit == "ug/l")
summary(ToPAFtwmicrogram$MeasuredValue) #values seem reasonable for oC -> assume unit is supposed to be oC
#Correct Tw measurements with unit in ug/l to unit oC
ToPAF$Unit[ToPAF$Parametercode == "Tw" & ToPAF$Unit == "ug/l"] <- "oC"
rm(ToPAFTw, ToPAFtwmicrogram)

#Tounit for Tw is oc, while Unit is oC
#Use of oC is actually consistent in Unit -> change ToUnit.
ToPAF$ToUnit[ToPAF$Parametercode == "Tw"] <- "oC"

# remove values of pH > 14
ToPAF <- ToPAF[-which(ToPAF$Parametercode == "pH" & ToPAF$MeasuredValue > 14), ]



#for n = 36 unit of PCDD48 measurements in ug/L (liter spelled with upper case instead of lower case)
#correct unit to ug/l
ToPAF$Unit[ToPAF$Parametercode == "PCDD48" & ToPAF$Unit == "ug/L"] <- "ug/l"

# n = 48 out of 557 cyffAd measurements in ul/l instead of ug/l
table(ToPAF$Unit[ToPAF$Parametercode == "cyffAd"])
cyffAd <- filter(ToPAF, Parametercode == "cyffAd")
# Taking a closer look at the measured values of both groups
# ggplot(cyffAd, aes(x = Unit, y = MeasuredValue)) + geom_boxplot()
summary(cyffAd$MeasuredValue[cyffAd$Unit == "ug/l"])  # Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 0.00200 0.00200 0.01000 0.01198 0.01000 0.05000
summary(cyffAd$MeasuredValue[cyffAd$Unit == "ul/l"])  # Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 0.00500 0.01000 0.01000 0.01042 0.01000 0.03000
rm(cyffAd)
#Assume manual input error in Unit -> correct name unit from ul/l to ug/l
ToPAF$Unit[ToPAF$Parametercode == "cyffAd" & ToPAF$Unit == "ul/l"] <- "ug/l"

#219 (out of 894972) chlorofyl A measurements in % instead of ug/l
table(ToPAF$Unit[ToPAF$Parametercode == "CHLFa"])
#219 measurements in % are removed from dataset
#create object defining which measurements need to be removed
CHLFa <- filter(ToPAF, Parametercode == "CHLFa")
CHLFaprct <- filter(CHLFa, Unit == "%")
#filter out measurements of CHLFa with unit %
ToPAF <- anti_join(ToPAF, CHLFa)
rm(CHLFa, CHLFaprct)

#6400 out of 394706 FEOa measurements with unit % instead of ug/l
table(ToPAF$Unit[ToPAF$Parametercode == "FEOa"])
#6400 measurements in % are removed from dataset
#create object defining which measurements need to be removed
FEOa <- filter(ToPAF, Parametercode == "FEOa")
FEOaprct <- filter(FEOa, Unit == "%")
#filter out measurements of minrlole with unit %
ToPAF <- anti_join(ToPAF, FEOaprct)
rm(FEOa, FEOaprct)

#Approximately half of all O2 measurements are in %, other half in concentration units.
#These are now normalized to "%" for saturation and "ug/l" for concentration.
table(ToPAF$Unit[ToPAF$Parametercode == "O2"])
O2temp <- filter(ToPAF, Parametercode == "O2" | Parametercode == "Tw" )

#4 out of 11619 minrole measurements with unit % instead of ug/l
table(ToPAF$Unit[ToPAF$Parametercode == "minrlole"])
#4 measurements in % are removed from dataset
#create object defining which measurements need to be removed
minrlole <- filter(ToPAF, Parametercode == "minrlole")
minrloleprct <- filter(minrlole, Unit == "%")
#filter out measurements of minrlole with unit %
ToPAF <- anti_join(ToPAF, minrloleprct)
rm(minrlole, minrloleprct)

save(ToPAF, file = "ToPAF.rda")
```

#### Check again for any remaining unit mismatches

```{r}
Unitmm <- unique(ToPAF[ToPAF$Unit!=ToPAF$ToUnit,c("Parametercode" ,"Unit", "ToUnit", "MolM")])
#O2 saturation is kept in % and O2 concentration is normalized to ug/l.
table(ToPAF$Unit[ToPAF$Parametercode == "O2"])

allowed_unit_mismatch <- rep(FALSE, nrow(ToPAF))
drop_unit_mismatch <- ToPAF$Unit != ToPAF$ToUnit & !allowed_unit_mismatch
drop_unit_mismatch[is.na(drop_unit_mismatch)] <- FALSE
if (any(drop_unit_mismatch)) {
  dropped_unit_rows <- drop_audit_subset(ToPAF[drop_unit_mismatch, , drop = FALSE])
  write_drop_audit(dropped_unit_rows, "Dropped_unresolved_unit_mismatches.csv")
  message("Dropping unresolved unit mismatches: ", nrow(dropped_unit_rows))
  ToPAF <- ToPAF[!drop_unit_mismatch, ]
}

drop_negative_values <- !is.na(ToPAF$MeasuredValue) &
  ToPAF$MeasuredValue < 0 &
  is_concentration_like_unit(ToPAF$Unit)
drop_negative_values[is.na(drop_negative_values)] <- FALSE
if (any(drop_negative_values)) {
  dropped_negative_rows <- drop_audit_subset(ToPAF[drop_negative_values, , drop = FALSE])
  write_drop_audit(dropped_negative_rows, "Dropped_negative_concentrations.csv")
  message("Dropping negative concentration-like measurements: ", nrow(dropped_negative_rows))
  ToPAF <- ToPAF[!drop_negative_values, ]
}

rm(Unitmm)
#Columns MolM, ToUnit, Keep and ValidCAS are no longer needed and thus removed
ToPAF$MolM <- NULL
ToPAF$ToUnit <- NULL
ToPAF$Keep <- NULL
ToPAF$ValidCAS <- NULL

# Remove measurements with NA "MeasuredValue" -----------------------------
ToPAF <- ToPAF[!is.na(ToPAF$MeasuredValue),]

# Change of names -----------------------------------------------------
#Change names of columns in "toPAF" df to the following:
names(ToPAF) <- c("SampleID", "AquoCode", "CAS", "MeasuredValue", "Unit", "Locationcode",
                  "AdditionalProperties", "LimitIndicator", "X", "Y", "DataSource", "SampleDate")

#translate CAS NVT to english
ToPAF$CAS[ToPAF$CAS == "NVT"] <- "Not Applicable"

```

#### Correction of some Measured Values based on expert judgement -----------------

hard change due of some Ca and Mg values due to very probable wrong unit, comm. Jaap P & Leo

```{r}
#Making a seperate save of the pre-adjusted Ca and Mg values
Ca_Mg <- filter(ToPAF, AquoCode %in% c("Ca", "Mg"))
write.csv(Ca_Mg, file = "Preadjusted_Data_Ca_Mg.csv", row.names = FALSE)
CaMgtest <- read.csv(file = "Preadjusted_Data_Ca_Mg.csv")

#n = 3164 Ca measurements changed from water board "Rijn en IJssel"
ToPAF$MeasuredValue[ToPAF$AquoCode=="Ca" & ToPAF$DataSource=="Rijn en IJssel" & ToPAF$MeasuredValue < 900] <-
  ToPAF$MeasuredValue[ToPAF$AquoCode=="Ca" & ToPAF$DataSource=="Rijn en IJssel" & ToPAF$MeasuredValue < 900] * 1000
#n = 3184 Mg measurements changed from water board "Rijn en IJssel"
ToPAF$MeasuredValue[ToPAF$AquoCode=="Mg" & ToPAF$DataSource=="Rijn en IJssel" & ToPAF$MeasuredValue < 95] <-
  ToPAF$MeasuredValue[ToPAF$AquoCode=="Mg" & ToPAF$DataSource=="Rijn en IJssel" & ToPAF$MeasuredValue < 95] * 1000
#n = 1054 Mg measurements changed from water board "Scheldestromen"
ToPAF$MeasuredValue[ToPAF$AquoCode=="Mg" & ToPAF$DataSource=="Scheldestromen" &
                      ToPAF$MeasuredValue < 1500 & ToPAF$SampleDate > "2005-01-01"] <-
  ToPAF$MeasuredValue[ToPAF$AquoCode=="Mg" & ToPAF$DataSource=="Scheldestromen" &
                        ToPAF$MeasuredValue < 1500 & ToPAF$SampleDate > "2005-01-01"] * 1000


# Harmonize input values "LimitIndicator" ---------------------------------
ToPAF$LimitIndicator[is.na(ToPAF$LimitIndicator)] <- "-"
ToPAF$LimitIndicator[ToPAF$LimitIndicator == " "] <- "-"

# Create unique SampleIDs based on unique date ("SampleDate") and location ("Meetpuntcode") combinations --------
#Facilitates analyzing mixtures at unique location-time combinations
#setup SampleIDs: unique date-locations
if (nrow(ToPAF) > 0) {
  ToSample <- ToPAF[, c("Locationcode", "SampleDate")]
  Astmp <- paste0(ToSample$Locationcode, ":", format(ToSample$SampleDate, "%Y-%m-%d"))
  ToSample$SampleID <- unlist(Astmp) #now it's not a list
  #Add SampleID to ToPAF dataframe
  ToPAF$SampleID[is.na(ToPAF$SampleID)] <- ToSample$SampleID
} else {
  ToPAF$SampleID <- character(0)
}
```

### Checkpoint 3: ToPAF.rds

```{r}
saveRDS(ToPAF, file = "ToPAF.rds")
```
