---
title: "0. Raw Input Processing Report"
author: "FOTO NL workflow"
date: "`r format(Sys.Date(), '%Y-%m-%d')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

if (!nzchar(Sys.getenv("SCRIPT_0_COLLECT_DEBUG", unset = ""))) {
  Sys.setenv(SCRIPT_0_COLLECT_DEBUG = "true")
}
if (!nzchar(Sys.getenv("DRY_RUN", unset = ""))) {
  Sys.setenv(DRY_RUN = "true")
}
if (!nzchar(Sys.getenv("SCRIPT_0_REPORT_PREVIEW_N", unset = ""))) {
  Sys.setenv(SCRIPT_0_REPORT_PREVIEW_N = "8")
}

resolve_script_0_report_dir <- function() {
  input_file <- tryCatch(knitr::current_input(dir = TRUE), error = function(...) "")
  if (nzchar(input_file) && file.exists(input_file)) {
    return(dirname(normalizePath(input_file)))
  }

  file_flag <- grep("^--file=", commandArgs(trailingOnly = FALSE), value = TRUE)
  if (length(file_flag) > 0L) {
    script_path <- sub("^--file=", "", file_flag[1])
    if (file.exists(script_path)) {
      return(dirname(normalizePath(script_path)))
    }
  }

  getwd()
}

script_dir <- resolve_script_0_report_dir()

source(file.path(script_dir, "helpers", "script_0", "build_input_common.R"))
source(file.path(script_dir, "helpers", "script_0", "build_input_board_configs.R"))

script_0_report_board <- Sys.getenv("SCRIPT_0_REPORT_BOARD", unset = "")
if (!nzchar(script_0_report_board)) {
  only_sources <- csv_arg(Sys.getenv("ONLY_SOURCES", unset = ""))
  if (length(only_sources) > 0L) script_0_report_board <- only_sources[[1]]
}
if (!nzchar(script_0_report_board)) {
  stop("Set SCRIPT_0_REPORT_BOARD or ONLY_SOURCES to render a board-level script_0 report.")
}

select_board_config <- function(configs, board_value) {
  board_key <- norm_key_text(board_value)
  matches <- Filter(function(cfg) {
    norm_key_text(cfg$datasource) == board_key || norm_key_text(cfg$board_id) == board_key
  }, configs)
  if (length(matches) == 0L) {
    stop("No board config matches: ", board_value)
  }
  matches[[1]]
}

ctx <- parse_context_args(commandArgs(trailingOnly = TRUE), base_dir = script_dir)
ctx$collect_debug <- TRUE
cfg <- select_board_config(get_board_configs(), script_0_report_board)
result <- run_board_reconstruction(cfg, ctx)
summary_dt <- result$summary
report <- result$report

fmt_runtime <- function(ctx) {
  data.frame(
    key = c(
      "datasource", "board_id", "raw_root", "reference_input_root", "output_root",
      "dry_run", "overwrite", "use_reference_filter", "collect_debug"
    ),
    value = c(
      report$datasource,
      report$board_id,
      ctx$raw_root,
      ctx$reference_input_root,
      ctx$output_root,
      ctx$dry_run,
      ctx$overwrite,
      ctx$use_reference_filter,
      ctx$collect_debug
    ),
    stringsAsFactors = FALSE
  )
}

missing_lookup <- function(stage_obj) {
  out <- stage_obj$missing
  if (nrow(out) == 0L) return(setNames(numeric(0), character(0)))
  vals <- out$missing_n
  names(vals) <- out$column
  vals
}

value_lookup <- function(stage_obj) {
  if (nrow(stage_obj$value_summary) == 0L) return(list(
    non_missing = NA_real_, negative_n = NA_real_, zero_n = NA_real_,
    min_value = NA_real_, max_value = NA_real_
  ))
  as.list(stage_obj$value_summary[1, , drop = FALSE])
}

stage_overview_table <- function(report) {
  stage_order <- c(
    "raw_combined",
    "mapped_template",
    "after_enrich",
    "after_board_logic",
    "after_generic_filter",
    "matched_reference",
    "final_output"
  )
  stage_order <- stage_order[stage_order %in% names(report$stages)]
  rows <- vector("list", length(stage_order))
  prev_rows <- NA_integer_
  for (i in seq_along(stage_order)) {
    nm <- stage_order[[i]]
    st <- report$stages[[nm]]
    vals <- value_lookup(st)
    miss <- missing_lookup(st)
    rows[[i]] <- data.frame(
      stage = nm,
      rows = st$n_rows,
      row_delta = if (is.na(prev_rows)) NA_integer_ else st$n_rows - prev_rows,
      cols = st$n_cols,
      meetwaarde_non_missing = vals$non_missing,
      negative_values = vals$negative_n,
      missing_date = unname(miss["Datum"]),
      missing_parametercode = unname(miss["Parametercode"]),
      missing_meetpunt = unname(miss["Meetpunt code"]),
      stringsAsFactors = FALSE
    )
    prev_rows <- st$n_rows
  }
  do.call(rbind, rows)
}

missing_delta_table <- function(before_stage, after_stage, label) {
  before <- missing_lookup(before_stage)
  after <- missing_lookup(after_stage)
  cols <- sort(unique(c(names(before), names(after))))
  if (length(cols) == 0L) return(data.frame())
  before_vals <- before[cols]
  after_vals <- after[cols]
  before_vals[is.na(before_vals)] <- 0
  after_vals[is.na(after_vals)] <- 0
  out <- data.frame(
    column = cols,
    missing_before = as.numeric(before_vals),
    missing_after = as.numeric(after_vals),
    stringsAsFactors = FALSE
  )
  out$filled_n <- pmax(out$missing_before - out$missing_after, 0)
  out$new_missing_n <- pmax(out$missing_after - out$missing_before, 0)
  out <- out[out$filled_n > 0 | out$new_missing_n > 0, , drop = FALSE]
  if (nrow(out) == 0L) return(out)
  out$change_context <- label
  out[order(-out$filled_n, -out$new_missing_n, out$column), c(
    "change_context", "column", "missing_before", "missing_after", "filled_n", "new_missing_n"
  )]
}

value_delta_table <- function(before_stage, after_stage, label) {
  before <- value_lookup(before_stage)
  after <- value_lookup(after_stage)
  data.frame(
    change_context = label,
    non_missing_before = before$non_missing,
    non_missing_after = after$non_missing,
    negative_before = before$negative_n,
    negative_after = after$negative_n,
    zero_before = before$zero_n,
    zero_after = after$zero_n,
    min_before = before$min_value,
    min_after = after$min_value,
    max_before = before$max_value,
    max_after = after$max_value,
    stringsAsFactors = FALSE
  )
}

compact_stage_snapshot <- function(stage_obj) {
  vals <- value_lookup(stage_obj)
  miss <- missing_lookup(stage_obj)
  data.frame(
    rows = stage_obj$n_rows,
    cols = stage_obj$n_cols,
    non_missing_meetwaarde = vals$non_missing,
    negative_values = vals$negative_n,
    missing_date = unname(miss["Datum"]),
    missing_parametercode = unname(miss["Parametercode"]),
    missing_meetpunt = unname(miss["Meetpunt code"]),
    stringsAsFactors = FALSE
  )
}
```

This report uses the same script_0 engine as the production path in `helpers/script_0/build_input_common.R`, but runs it in `collect_debug` mode so each major transformation step is summarized.

## Runtime Context

```{r runtime-context}
knitr::kable(fmt_runtime(ctx), format = "simple")
```

## Board Config

```{r board-config}
knitr::kable(data.frame(
  field = c("datasource", "board_id", "output_file", "raw_patterns_n"),
  value = c(cfg$datasource, cfg$board_id, cfg$output_file, length(cfg$raw_patterns)),
  stringsAsFactors = FALSE
), format = "simple")
cat("\nRaw patterns:\n")
print(cfg$raw_patterns)
```

## Raw File Discovery

```{r raw-file-discovery}
knitr::kable(report$raw_file_summary, format = "simple")
```

## Stage Overview

```{r stage-overview}
knitr::kable(stage_overview_table(report), format = "simple")
```

## Enrichment Impact

```{r stage-enrichment-impact}
enrich_delta <- missing_delta_table(report$stages$mapped_template, report$stages$after_enrich, "mapped_to_enriched")
if (nrow(enrich_delta) == 0L) {
  cat("No field-level missingness changes detected during enrichment.\n")
} else {
  knitr::kable(enrich_delta, format = "simple")
}
```

## Board Logic Impact

```{r stage-board-logic-impact}
board_logic_delta <- missing_delta_table(report$stages$after_enrich, report$stages$after_board_logic, "enriched_to_board_logic")
board_logic_values <- value_delta_table(report$stages$after_enrich, report$stages$after_board_logic, "enriched_to_board_logic")
if (nrow(board_logic_delta) == 0L) {
  cat("No field-level missingness changes detected during board logic.\n\n")
} else {
  print(knitr::kable(board_logic_delta, format = "simple"))
  cat("\n")
}
print(knitr::kable(board_logic_values, format = "simple", digits = 3))
```

## Generic Filter Impact

```{r stage-filter-impact}
filter_delta <- missing_delta_table(report$stages$after_board_logic, report$stages$after_generic_filter, "board_logic_to_generic_filter")
filter_values <- value_delta_table(report$stages$after_board_logic, report$stages$after_generic_filter, "board_logic_to_generic_filter")
filter_summary <- data.frame(
  rows_before = report$stages$after_board_logic$n_rows,
  rows_after = report$stages$after_generic_filter$n_rows,
  rows_dropped = report$stages$after_board_logic$n_rows - report$stages$after_generic_filter$n_rows,
  stringsAsFactors = FALSE
)
print(knitr::kable(filter_summary, format = "simple"))
cat("\n")
if (nrow(filter_delta) > 0L) {
  print(knitr::kable(filter_delta, format = "simple"))
  cat("\n")
}
print(knitr::kable(filter_values, format = "simple", digits = 3))
```

## Reference Match

```{r reference-comparison}
knitr::kable(data.frame(
  field = c("reference_file", "reference_rows", "matched_rows", "coverage", "signature_mode", "sig_intersect_unique"),
  value = c(
    report$reference$reference_file,
    report$reference$reference_rows,
    report$reference$matched_rows,
    report$reference$coverage,
    report$reference$signature_mode,
    report$reference$sig_intersect_unique
  ),
  stringsAsFactors = FALSE
), format = "simple")
if (length(report$reference$read_errors) > 0L) {
  cat("\nRead errors:\n")
  print(report$reference$read_errors)
}
```

## Final Output Snapshot

```{r final-output}
knitr::kable(compact_stage_snapshot(report$stages$final_output), format = "simple", digits = 3)
cat("\nSample rows:\n")
print(knitr::kable(report$stages$final_output$sample, format = "simple"))
```

## Diagnostics And Output

```{r diagnostics-output}
knitr::kable(data.frame(
  field = c("board_out_dir", "output_file", "diagnostic_file", "read_error_file"),
  value = c(
    report$diagnostics$board_out_dir,
    report$diagnostics$output_file,
    report$diagnostics$diagnostic_file,
    report$diagnostics$read_error_file
  ),
  stringsAsFactors = FALSE
), format = "simple")
cat("\nSummary row:\n")
print(knitr::kable(report$diagnostics$summary, format = "simple"))
```
